---
title: "Exploratory Data Analysis"
output: 
  pdf_document:
    fig_caption: yes
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(gridExtra)
library(kableExtra)
data <- read_csv("../cleaned_data.csv")
```  


```{r, fig.width = 7}
kable(summary(data)) %>% kable_styling()
```

```{r OS vs Stress, fig.height = 3, fig.width = 7, fig.caption = "\\label{fig:figs} Distribution of Stress Level by Operating System"}

data %>% 
  filter(os != "Linux") %>% 
  ggplot(aes(x=os, y=stress)) +
  geom_boxplot() +
  theme_bw() +
  ylab("Stress Level") +
  xlab("Operating System") +
  ggtitle("Operating System vs Stress Level") +
  theme(axis.title.x = element_text(margin = margin(t = 15, r = 0, b = 10, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 10)))
```  
  
It's interesting to note that the median stress level is not really different between the two operating systems, but the distribution of stress actually is: Figure 1 shows that most Windows users are concentrated around higher stress levels, whereas a significant part of Mac users present mid-low levels of stress (< 7 in the 0-10 scale).  

```{r coding plots,  fig.width = 7, fig.height = 4, fig.caption = "Coding Skills"}
data$coding <- fct_relevel(data$coding, 
                           "high",
                           "moderate",
                           "average",
                           "low",
                           "none")
coding_stress <- data %>% 
  ggplot(aes(y = stress, x = coding)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5) +
  ggtitle("Coding Skill vs. Stress Level") +
  xlab("coding skill") +
  theme_bw()


coding_os <- data %>% 
  filter(os != "Linux") %>% 
  ggplot(aes(x = coding)) +
  geom_bar(aes(fill = os )) +
  scale_fill_viridis_d(name = "") +
  ggtitle("Coding Skill vs. Operating System") +
  xlab("coding skill") +
  theme_bw() 

grid.arrange(coding_stress, coding_os, nrow = 2)
```    
```{r experience plots, fig.height = 3, fig.width = 7, echo = FALSE}
experience_stress <- data %>%
  filter(work_year < 11) %>% 
  ggplot(aes(x = work_year, y = stress)) +
  geom_jitter(alpha = 0.5) +
  theme_bw() +
  xlab("years of work experience") + 
  ggtitle("Experience vs. Stress Level")

experience_os <- data %>% 
  filter(os != "Linux", work_year < 11) %>%
  ggplot(aes(y = work_year, x = os)) +
  geom_boxplot()+
  geom_jitter(alpha = 0.5) +
  theme_bw() +
  ylab("years of work experience") + 
  xlab("") +
  ggtitle("Experience vs. Operating System") 

grid.arrange(experience_stress, experience_os, ncol = 2)
```  


```{r gender plots, fig.height = 3, fig.width = 7,  fig.caption = "Gender"}

gender_stress <- data %>% 
  ggplot(aes(y = stress, x = gender)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5) +
  theme_bw() +
  ggtitle("Gender vs. Stress Level") +
  xlab("")

gender_os <- data %>% 
  filter(os != "Linux") %>% 
  ggplot(aes(gender)) +
  geom_bar(aes(fill = os)) +
  scale_fill_viridis_d(name = "") +
  theme_bw() +
  ggtitle("Gender vs. Operating System") +
  xlab("")

grid.arrange(gender_stress, gender_os, ncol = 2)
  

```    

```{r background plots, fig.height = 4, fig.width = 7, fig.caption = "Background"}

background_stress <- data %>%
  filter(background != "Arts, Humanities, and Social Sciences") %>% 
  ggplot(aes(x = background, y = stress)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5) +
  xlab("") +
  coord_flip() +
  theme_bw() +
  ggtitle("Background vs. Stress Level")

background_os <- data %>% 
  filter(os != "Linux",
         background != "Arts, Humanities, and Social Sciences") %>% 
  ggplot(aes(x = background)) +
  geom_bar(aes(fill = os )) +
  scale_fill_viridis_d(name = "") +
  ggtitle("Background vs. Operating System") +
  xlab("") +
  coord_flip() +
  theme_bw() 
grid.arrange(background_stress, background_os, nrow = 2)
```  






